---
title: "Descriptive Statistics"
author: "Miranda Rintoul"
date: "9/7/2022"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Histograms

Let's load the milk fat data and build a histogram.

```{r}
# "c" builds a vector of data points
milkfat <- c(20, 38, 44, 49, 42, 21, 48, 47, 21,
           22, 23, 26, 26, 44, 48, 29)

sort(milkfat)
sort(milkfat, decreasing = TRUE)
```

Most basic histogram:

```{r}
hist(milkfat)
```

We can set arguments to the `hist` function to add more features.

```{r}
# "mean" for title, "col" for bar colors
hist(milkfat, main = "Histogram of Milk Fat Data",
     col = "dodgerblue")
```

Setting breaks by hand with the `seq()` function:

```{r}
# Bins of width 5, from 20 to 50
hist(milkfat, main = "Histogram of Milk Fat Data",
     col = "dodgerblue", breaks = seq(20, 50, 5))

# Bins of width 7, from 14 to 49
hist(milkfat, main = "Histogram of Milk Fat Data",
     col = "dodgerblue", breaks = seq(14, 49, 7))
```

We can also set the `breaks` option to be a single number.

```{r}
# Many breaks = narrow bins
hist(milkfat, main = "Histogram of Milk Fat Data",
     col = "dodgerblue", breaks = 50)

# Few breaks = wide bins
hist(milkfat, main = "Histogram of Milk Fat Data",
     col = "dodgerblue", breaks = 4)
```


## Location

Let's get the median and mean of the milk fat data.

```{r}
median(milkfat)

# "Long hand" mean calculation
sum(milkfat) / length(milkfat)

# R shortcut mean calculation
mean(milkfat)
```

To get the percentiles/quantiles of the data, we need the ``quantile`` function.  By default, this gives the quartiles.

```{r}
quantile(milkfat)
```

To find a specific percentile, specify the ``probs`` argument to be a vector of probabilities.  For example, let's get the 32nd, 58th, and 97th percentiles.

```{r}
quantile(milkfat, probs = c(0.32, 0.58, 0.97))
```

What if the value 49 was accidentally recorded as 94?

```{r}
milkfat_error <- c(20, 38, 44, 94, 42, 21, 48, 47, 21,
           22, 23, 26, 26, 44, 48, 29)

# Find mean and median of new error data
```


## Spread

Let's begin by finding the IQR of the original milkfat data.

```{r}
IQR(milkfat)
```

This is consistent with our previous output:  we saw that the 25th percentile of the data is 22.75, and the 75th percentile is 44.75.  22.75 - 44.75 = 22.

The ``range`` function in R will output the minimum and maximum, by default.  So, we need to use subtraction to find the range.

```{r}
range(milkfat)
```

The range is 49 - 20 = 29.  

There are a few different ways we can find the standard deviation. Long way:

```{r}
mean(milkfat)

# Deviations from mean
milkfat - mean(milkfat)

# Sum of squared deviations
sum((milkfat - mean(milkfat))^2) 

# Sample sd via formula
sqrt(sum((milkfat - mean(milkfat))^2)/15) 
sqrt(sum((milkfat - mean(milkfat))^2)/(length(milkfat)-1))

```

Short way:

```{r}
# Sample sd shortcut
sd(milkfat)

# Sample variance shortcut
var(milkfat)
```

Finally, let's see how these measures of spread change when we look at the data with an error.

```{r}
# Compare IQR, range, and SD for both datasets

```


## Other graphs

Let's make a very basic boxplot, then improve on it.

```{r}
boxplot(milkfat)

boxplot(milkfat, horizontal = TRUE, xlab = "Milk Fat")
```

Here is an example with multiple boxplots:

```{r}
A <- c(67, 78, 88, 65, 77, 72, 73, 82, 85, 90)
B <- c(77, 61, 85, 75, 77, 73, 81, 74, 83, 89)
C <- c(84, 82, 79, 76, 88, 90, 91, 85, 82, 94)

boxplot(A, B, C, main = "Exam Scores by Program",
        col = c("#D81B60", "#1E88E5", "#FFC107"),
        names = c("A", "B", "C"))
```


## Multiple datasets

We have data on milk fat for cows with and without supplements.

```{r}
milkfat_supp <- milkfat
milkfat_nosupp <- c(20, 21, 22, 22, 23, 24, 18, 27, 25,
                    25, 19, 21, 28, 18, 17, 21, 23)
```

Let's build boxplots with default axes.

```{r}
par(mfrow = c(1, 2))  # View 2 plots at once
hist(milkfat_supp, main = "Supplement Milk Fat",
     col = "dodgerblue")

hist(milkfat_nosupp, main = "No Supplement Milk Fat",
     col = "dodgerblue")
```

Now, let's fix the axes so that the x and y axes (and bins) are consistent.

```{r}
par(mfrow = c(1, 2))  # View 2 plots at once
hist(milkfat_supp, main = "Supplement Milk Fat",
     col = "dodgerblue", ylim = c(1, 8),
     breaks = seq(16, 49, 3))

hist(milkfat_nosupp, main = "No Supplement Milk Fat",
     col = "dodgerblue", ylim = c(1, 8),
     breaks = seq(16, 49, 3))
```

We can build a boxplot as well.

```{r}
boxplot(milkfat_supp, milkfat_nosupp, main = "Milk Fat",
        names = c("Supplement", "No Supplement"))
```

Finally, we can compare numeric summaries.

```{r}
mean(milkfat_supp); mean(milkfat_nosupp)
median(milkfat_supp); median(milkfat_nosupp)
sd(milkfat_supp); sd(milkfat_nosupp)
```

We can combine the vectors easily in R if we want to summarize the entire group of cows at once.

```{r}
combined <- c(milkfat_supp, milkfat_nosupp)
combined
```

From here we can create all of the usual summaries (histogram, mean, sd, etc.) for a single set of data.


## New R Functions

c() - combine several values into a data vector
sort() - sort a data vector into ascending (or descending) order
hist() - create a histogram of a data vector

median() - find the median value of a data vector
sum() - add all of the values in a data vector
length() - find the length (number of values) of a data vector
mean() - find the average value of a data vector
quantile() - find any quantile/percentile of a data vector

IQR() - find the interquartile range of a data vector
range() - find the range of a data vector
sqrt() - find the square root of a number
sd() - find the standard deviation of a data vector
var() - find the variance of a data vector

boxplot() - create a boxplot of one or more data vectors

