---
title: "Random Variables"
author: "Miranda Rintoul"
date: "9/14/2022"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Binomial RVs

In the genes example, X is a binomial RV with n = 3 and pi = 9/16.  What is the probability that X = 2?

```{r}
dbinom(2, size = 3, prob = 9/16)
```

Genes example: find the probability that X = 1 and X \>= 1.

```{r}
# First part: P(X = 1)
dbinom(1, size = , prob = )

# Different ways to do the second problem:
1 - dbinom(0, size = 3, prob = 9/16) # subtraction
1:3
dbinom(1:3, size = 3, prob = 9/16)
sum(dbinom(1:3, size = 3, prob = 9/16)) # add probabilities for 1-3
```

## Normal RVs

The height of US males is normally distributed with mean 70 inches and standard deviation 3 inches.  What is the probability that a randomly chosen US male is below 72 inches?

```{r}
# With standardization:
pnorm(2/3, mean = 0, sd = 1)
pnorm(2/3)    # by default, pnorm is N(0, 1)

# Directly:
pnorm(72, mean = 70, sd = 3)
```

Package weights have a normal distribution with mean 16.1 ounces and sd 0.3 ounces. What is the probability that a randomly chosen package weighs less than 16 ounces?

```{r}

```

What is the probability that a randomly chosen package has weight between 15.5 and 16.5 ounces?

```{r}

```

Above what weight is the heaviest 1% of packages?

```{r}
qnorm(0.99, 16.1, 0.3) # units of W

# Or, find a standardized score and convert back to W
qnorm(0.99, 0, 1)
qnorm(0.99, 0, 1)*0.3 + 16.1
```

What package weight is the 3rd quartile?

```{r}

```

Suppose ten packages are randomly selected off the production line. What is the probability that exactly 6 packages have weights above 16 ounces?

First, find the probability of a single package being above 16 ounces.

```{r}
1 - pnorm(16, mean = 16.1, sd = 0.3)
pnorm(16, mean = 16.1, sd = 0.3, lower.tail = F)
```

If Y is the number of packages above 16 ounces, then Y is binomial with n = 10 and pi = 0.6306.  We can use a binomial probabilty to find the probability that exactly 6 packages have weights above 16 ounces.

```{r}
dbinom()
```

## New R Functions

dbinom() - find a specific probability of a binomial distribution

pnorm() - find the area (probability) to the left of a value on a normal distribution
qnorm() - find a specific quantile/percentile of a normal distribution





