---
title: "Sampling Distributions and Estimation"
author: "Miranda Rintoul"
date: "9/26/2022"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Distribution of X-bar

Broiler weights have a N(1.397, 0.161^2) distribution.  What is the probability that a randomly selected broiler has weight below 1.2?

```{r}
pnorm(1.2, 1.387, 0.161)
```

If we sample 5 broilers, what is the probability that the average weight is below 1.2?

```{r}
pnorm()
```

If we sample 50 broilers, what is the probability that the average weight is below 1.2?

```{r}
pnorm()
```

Now consider the sample total. What is the 95th percentile for the total weight of 50 broilers?

```{r}
qnorm()
```

## CLT

```{r}
# "x" is 15 points of normal data
x <- rnorm(15, 0, 1)

qqnorm(x)
```

Using the CLT, we can approximate the mean number of moths in 100 traps as N(0.5, (0.7^2)/100).  What is the probability that the mean number of moths in 100 traps is above 0.75?

```{r}
pnorm()
```

Suppose we have a population proportion pi = 0.07.  Then a sample proportion from a sample of 300 should have a normal distribution with mean 0.07 and variance (0.07*0.93)/300.

What is the probability that p_hat is less than 8/300?

```{r}
pnorm()
```

## New R Functions

qqnorm() - create a normal qq-plot of a data vector
rnorm() - draw a random sample from a normal distribution






